---
title: "Build a phyloseq object"
output: html_notebook
---

Let's start by reading your data:

```{r}
otus <- read.csv("genera1.csv", check.names=F)
taxa <- read.csv("Taxa-1.csv", sep="\t", check.names=F)

otus <- otus[!duplicated(otus$public_client_id), ]
genus_cols <- grepl("genus_", names(otus))
rownames(otus) <- otus$public_client_id

sdata <- otus[, !genus_cols]
alleles <- paste0(sdata$rs429358, sdata$rs7412)
sdata$apoe <- "healthy"
sdata$apoe[alleles %in% c("CTCT", "CCCT", "CTCC")] <- "heterozygous Apo-e4"
sdata$apoe[alleles == "CCCC"] <- "homozygous Apo-e4"
sdata$apoe <- factor(sdata$apoe, levels=c("healthy", "heterozygous Apo-e4", "homozygous Apo-e4"))
table(sdata$apoe)

otus <- as.matrix(otus[, genus_cols])
colnames(otus) <- gsub("genus_", "", colnames(otus))
```

Now we convert it to fit in a phyloseq object. We need to fulfill the following rules:

1. The OTU table is a matrix with rows being samples and each column being a taxon.
2. The taxonomy table must be a matrix with rows being taxa and columns being ranks.
3. The sample data must be data frame with rows being samples.

```{r}
library(phyloseq)

tax_matrix <- as.matrix(taxa[, 2:ncol(taxa)])
rownames(tax_matrix) <- taxa[, 1]
```

To double check if everything works let's do some validity checks:

```{r}
stopifnot(all(rownames(otus) %in% rownames(sdata)))
stopifnot(nrow(otus) == nrow(sdata))
print("sample names match")

stopifnot(all(colnames(otus) %in% rownames(tax_matrix)))
stopifnot(ncol(otus) == nrow(tax_matrix))
stopifnot(!anyDuplicated(tax_matrix))
print("taxa look okay")

stopifnot(!anyDuplicated(sdata))
print("sample data looks okay")
```

If that passes we can go ahead and build our phyloseq object.

```{r}
ps <- phyloseq(
  otu_table(otus, taxa_are_rows = FALSE),
  tax_table(tax_matrix),
  sample_data(sdata)
)

ps
```